\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)

### 2015-09-03 background.tex
some updates
updates in overview
addressed most of Shuo's comments

Easy to think you can do better until it happens to you :-)

\subsection{Programming and testing with \psharp}
\label{bg:psharp}

The \psharp~\cite{deligiannis2015psharp} project provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} framework for developing highly-reliable distributed systems.

The \psharp language is an extension of \csharp that enables asynchronous programming using communicating state-machines. \psharp machines can interact asynchronously by sending and receiving events, an approach commonly used to develop distributed systems. This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g. Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).

A \psharp machine consists of an input queue for received events, states, state transitions, event handlers, fields and methods. Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler. This handler might update a field, create a new machine, or send an event to another machine. In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.

Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code. Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.

The second core component of the \psharp project is a concurrency unit testing framework that is embedded in the \psharp runtime. This framework allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.

\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.

>>>>

Distributed systems typically consist of two or more components that communicate \emph{asynchronously} by sending and receiving messages through a network layer~\cite{lamport1978time}. Each component has its own input message queue, and when a message arrives, the component responds by executing an appropriate \emph{message handler}. Such a handler consists of a sequence of program statements that might update the internal state of the component, send a message to another component in the system, or even create an entirely new component.

\subsection{Challenges in testing}
\label{sec:bg:challenges}

In a distributed system, message handlers can interleave in arbitrary order, because of the asynchronous nature of message-based communication. To complicate matters further, unexpected failures are the norm in production systems: nodes in a cluster might fail at any moment, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and recover the state of the system. Moreover, with multicore machines having become a commodity, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.

All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques. Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code. However, these techniques are not effective for testing distributed systems, as they are not able to capture and control the many sources of nondeterminism.

The ideal testing technique should be able to work on unmodified distributed systems, capture and control all possible sources of nondeterminism, systematically inject faults in the right places, and explore all feasible execution paths. However, this is easier said than done when testing production systems.

A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.  A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking. Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}. A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system. Even if the model is verified, the gap between a real-world implementation and the verified model is still significant, so implementation bugs are still a realistic concern.

\subsection{Types of bugs}
\label{sec:bg:bugs}

We can classify most distributed system bugs in two categories: \emph{safety} and \emph{liveness} property violations~\cite{lamport1977proving}.

\begin{description}
\item[Safety] A safety property checks that an erroneous program state is \emph{never} reached, and is satisfied if it \emph{always} holds in each possible program execution.

\item[Liveness] A liveness property checks that some progress \emph{will} happen, and is satisfied if it \emph{always eventually} holds in each possible program execution.
\end{description}

\noindent
A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state. An example of a generic safety property for message passing systems is to assert that whenever a message gets dequeued there must be an action that can handle the received message.

Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states. Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}. Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property. Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.

### 2015-09-03 casestudies.tex
addressed most of Shuo's comments

% N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.  As the traffic continues to grow over time, we will need \SCComment{change to "the system needs", because the word "we" is unclear} to reshard the data set across a greater number of Azure storage accounts without interrupting service.  During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration (Figure~\ref{fig:livemigration}).  MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.

>>>>

% N.B. Artifact Services is mentioned at http://research.microsoft.com/en-us/people/schulte/.  Hopefully it's OK to reveal that it was the system in this case study. ~ Matt 2015-08-17
The initial motivation for MigratingTable was to solve a scaling problem for Artifact Services, an internal Microsoft system with a data set that is sharded across tables in different Azure storage accounts because it exceeds the limit on traffic supported by a single Azure storage account.  As the traffic continues to grow over time, the system needs to reshard the data set across a greater number of Azure storage accounts without interrupting service.  During such a resharding, our sharding manager will identify each key range that should migrate to a different table, and we will use a separate MigratingTable instance for each such key range to actually perform the migration (Figure~\ref{fig:livemigration}).  MigratingTable may also be useful to migrate data to a table with different values of configuration parameters that Azure does not support changing on an existing table, such as geographic location.

### 2015-09-03 intro.tex
fixed compiler errors
addressed most of Shuo's comments

Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}. This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
%
All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute. A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}.\SCComment{I suggest we add "but the consequence can be catastrophic". Otherwise, why does such an extremely unlikely bug matter?}

Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism \SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?}, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production. Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money and many dissatisfied customers. \SCComment{do we have a few real stories about the damages? They are helpful.}

We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development, and the unified response \SCComment{change "unified response" to "consensus"?} was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production. The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, publicly acknowledge~\cite{newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems.

Amazon recently published an article~\cite{newcombe2015aws} that describes their use of TLA+~\cite{lamport1994temporal} to detect distributed system bugs and prevent them from reaching production. TLA+ is a powerful specification language for verifying distributed protocols, but it is unable to verify the code that is actually being executed. The implied assumption is that a model of the system will be verified, and then the programmers are responsible to match what was verified with the source code of the real system. Although many design bugs can be caught with this approach, there is \emph{no guarantee} that the real distributed system will be free of bugs.\SCComment{we want to avoid making a general point that TLA+ has no guarantee to find all bugs, because our approach cannot either. Instead, we should say something like "as we will show in the paper, many bugs we found are too subtle to be discovered without examining concrete implementations."}

@@

\SCComment{let me give my overall comments about the current version of the paper here. This set of case studies are really cool. The fact that our approach is proven to be effective clearly is the primary selling point of the paper. Reviewers will love the results we have.

About the writing, there may be 3 areas to improve:
(1) the word "we" is overloaded, because "we" play multiple roles in this work. "We" built \psharp; "we", as developers, designed the TableMigration system; "we", as testers, also take the responsibility to test the legacy Azure vNext system. The current writing doesn't clearly differentiate these roles, so readers may wonder how much work does it requires if they want to adopt the approach. 
(2) The intro makes me feel that we compare with existing logic reasoning approaches such as TLA+, and point out the limitation of TLA+. Until much later in the paper, I realize that our approach doesn't have any logic reasoning ingredient in it. I suggest that the intro makes it clear that our approach is complementary to TLA+, not a competing approach. We should avoid give a comparison with TLA+, because it is not an apple-to-apple comparison.
(3) Section 4 is the most important section about the approach. I feel that we spend too much text on section 4.3, which is about building the test harness. My concern is that readers' minds may be dominated by how we mocked all kinds of objects and override real methods by fake methods. They may feel that, in the end, we tested a highly simplified system. This is not true, because the extent manager is a REALLY complicated system (which is shown as a small box in fig.3). I suggest that we zoom into this system to explain why many things can go wrong and it is hard to test in a conventional manner.}

>>>>

Distributed systems are notoriously hard to design, implement and test~\cite{cavage2013there, laguna2015debugging, maddox2015test}. This is due to many well-known sources of \emph{nondeterminism}~\cite{chandra2007paxos}, such as race conditions in the asynchronous interaction between system components, the use of multithreaded code inside a component, unexpected node failures, unreliable communication channels and data losses, and interaction with (human) clients.
%
All these sources of nondeterminism translate into \emph{exponentially} many execution paths that a distributed system might potentially execute. A bug might hide deep inside one of these paths and only manifest under extremely rare corner cases~\cite{gray1986computers, musuvathi2008finding}, but the consequence can be catastrophic.

Classic techniques that product groups employ to test, verify and validate their systems (such as code reviews, unit testing, stress testing and fault injection) are unable to capture and control all the aforementioned sources of nondeterminism \SCComment{my comment is similar to the one in the intro: do we claim that we capture all sources of nondet?} \PDComment{only the ones that the user has mocked using the \psharp APIs, some are handled automatically like async/await, we should probably rephrase}, which causes the most tricky bugs being missed during testing and only getting exposed after a system has been put in production. Discovering and fixing bugs in production, though, is bad for business as it can cost a lot of money and many dissatisfied customers. \SCComment{do we have a few real stories about the damages? They are helpful.}

We interviewed engineers from the Microsoft Azure team regarding the top problems in distributed system development, and the consensus was that the most critical problem today is how to improve \emph{testing coverage} to find bugs \emph{before} a system goes in production. The need for better testing techniques is not specific to Microsoft; other companies, such as Amazon and Google, publicly acknowledge~\cite{newcombe2015aws} that testing methodologies have to improve to be able to reason about the correctness of increasingly more complex distributed systems. \PDComment{We are missing a link for Google, but would be good if we can provide such a link, or for other companies like Facebook, I will try find something, but if anyone has a good suggestion please include}

Amazon recently published an article~\cite{newcombe2015aws} that describes their use of TLA+~\cite{lamport1994temporal} to detect distributed system bugs and prevent them from reaching production. TLA+ is a powerful specification language for verifying distributed protocols, but it is unable to verify the code that is actually being executed. The implied assumption is that a model of the system will be verified, and then the programmers are responsible to match what was verified with the source code of the real system. Although many design bugs can be caught with this approach, there is \emph{no guarantee} that the real distributed system will be free of bugs.\SCComment{we want to avoid making a general point that TLA+ has no guarantee to find all bugs, because our approach cannot either. Instead, we should say something like "as we will show in the paper, many bugs we found are too subtle to be discovered without examining concrete implementations."} \PDComment{We should rephrase. By the way the main contrast with TLA+ is that TLA+ works \emph{only} on a model and not the real code, whereas \psharp works on a real code (with a modeled environment).}

@@

\SCComment{let me give my overall comments about the current version of the paper here. This set of case studies are really cool. The fact that our approach is proven to be effective clearly is the primary selling point of the paper. Reviewers will love the results we have.\\
About the writing, there may be 3 areas to improve:
(1) the word "we" is overloaded, because "we" play multiple roles in this work. "We" built \psharp; "we", as developers, designed the TableMigration system; "we", as testers, also take the responsibility to test the legacy Azure vNext system. The current writing doesn't clearly differentiate these roles, so readers may wonder how much work does it requires if they want to adopt the approach. 
(2) The intro makes me feel that we compare with existing logic reasoning approaches such as TLA+, and point out the limitation of TLA+. Until much later in the paper, I realize that our approach doesn't have any logic reasoning ingredient in it. I suggest that the intro makes it clear that our approach is complementary to TLA+, not a competing approach. We should avoid give a comparison with TLA+, because it is not an apple-to-apple comparison.
(3) Section 4 is the most important section about the approach. I feel that we spend too much text on section 4.3, which is about building the test harness. My concern is that readers' minds may be dominated by how we mocked all kinds of objects and override real methods by fake methods. They may feel that, in the end, we tested a highly simplified system. This is not true, because the extent manager is a REALLY complicated system (which is shown as a small box in fig.3). I suggest that we zoom into this system to explain why many things can go wrong and it is hard to test in a conventional manner.}  \PDComment{Really nice suggestions. I totally agree with the last point you make and we should make sure this is addressed (it is super important)!}

### 2015-09-03 methodology.tex
addressed most of Shuo's comments

\item The top level asynchrony due to message passing, has to be exposed as event sending using the \psharp runtime API. If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing. This step allows the \psharp runtime to capture and control the message passing interleavings during systematic testing.

\item Any source of nondeterminism in the system or the environment (e.g. failures and timers), must be modeled using the \psharp \texttt{Nondet()} method. This method returns a nondeterministic boolean value \SCComment{why only Boolean? even a timer returns a Boolean value?} that is controlled by the \psharp runtime and can be used to systematically inject ``nondeterministic'' events and for bug reproduction.
\end{enumerate}

@@

The \texttt{Environment} is essentially a ``god'' machine: \SCComment{it is helpful to indicate this machine clearly in Fig.3. I think it is the dashed line box} it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute (e.g. how many nodes should be created, how many failures should occur \SCComment{why is the failure rate limited? is there a supporting evidence for the limit we set?}, etc).

@@

Our methodology in \psharp to systematically test distributed systems that rely on timers, is to abstract timers away, model them using message passing communication and introduce nondeterminism in their firing.

Figure~\ref{fig:timer} shows how we modeled a generic timer in the Azure Storage vNext case study. The Extend Manager, as well as each Extent Node in the harness, is associated with a unique \texttt{Timer} machine. When creating this machine, we pass as a payload the id of the machine that owns this timer. When the \texttt{Timer} machine is created, it stores this id in the \texttt{Owner} field and then transitions to the \texttt{Active} state. In this state, the \texttt{Timer} loops infinitely and nondeterministically sends a \texttt{TimerTickEvent} to \texttt{this.Owner}. When the Extent Node owner receives this event, it handles it by generating a synchronization message that is being send to the Extent Manager. Similarly, when the Extent Manager receives a \texttt{TimerTickEvent} from its own \texttt{Timer}, it handles it by nondeterministically invoking repair-related methods in the Extent Repair Center data structure.

>>>>

\item The top level asynchrony due to message passing, must be exposed as event sending using the \psharp APIs. If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing. This step allows the \psharp runtime to capture and control the nondeterminism due to message passing interleavings during systematic testing.

\item Any other source of nondeterminism in the system or the environment (e.g. failures and timers), must be explicitly modeled using the available \psharp APIs. This would allow \psharp to capture the nondeterminism and control it during systematic testing.
\end{enumerate}

@@

The \texttt{Environment} (denoted with a dashed line box in Figure~\ref{fig:azurestoremodel}) is essentially a ``god'' machine: it is the very first machine that is created; has a handle to all other machines in the harness; and includes the logic based on which the test harness will execute. The programmer is free to create either a finite test harness (e.g. with a predefined number of node failures) or an infinite test harness (e.g. with an unbounded number of failures), as long as any nondeterminism (e.g. when to insert a failure) is captured using the available \psharp APIs.

@@

Our methodology in \psharp to systematically test distributed systems that rely on timers, is to abstract timers away, model them using message passing communication and introduce nondeterminism in their firing. This nondeterminism is introduced using the \psharp \texttt{Nondet()} method, which returns a nondeterministic boolean value that is controlled by the \psharp runtime during systematic testing.

Figure~\ref{fig:timer} shows how we modeled a generic timer in the Azure Storage vNext case study. The Extend Manager, as well as each Extent Node in the harness, is associated with a unique \texttt{Timer} machine. When creating this machine, we pass as a payload the id of the machine that owns this timer. When the \texttt{Timer} machine is created, it stores this id in the \texttt{Owner} field and then transitions to the \texttt{Active} state. In this state, the \texttt{Timer} loops infinitely and nondeterministically (using \texttt{Nondet()}) sends a \texttt{TimerTickEvent} to \texttt{this.Owner}. When the Extent Node owner receives this event, it handles it by generating a synchronization message that is being send to the Extent Manager. Similarly, when the Extent Manager receives a \texttt{TimerTickEvent} from its own \texttt{Timer}, it handles it by nondeterministically invoking repair-related methods in the Extent Repair Center data structure.

### 2015-09-03 paper.tex
addressed most of Shuo's comments

We present a new methodology for systematically testing distributed systems. Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism\SCComment{this paper mainly focuses on the nondet due to scheduling, right? There are other sources, e.g., arbitrary input, system call behavior, clock, random number, etc. I feel that ``all sources'' is perhaps an overclaim}. We present two case studies of using \psharp to test production distributed systems inside Microsoft. Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it. \psharp uncovered the bug in a very small setting, which made it easy to examine traces, identify and fix the problem. \PDComment{we now have many bugs in Matt's system too.}

@@

\section{Distributed Systems \SCComment{change to "Distributed System Bugs"?}}
\label{sec:bg}

\input{overview}

>>>>

We present a new methodology for systematically testing distributed systems. Our approach involves the use of \psharp, an extension of \csharp that combines a flexible environmental modeling approach with a concurrency testing framework, which can capture and systematically explore all sources of nondeterminism \SCComment{this paper mainly focuses on the nondet due to scheduling, right? There are other sources, e.g., arbitrary input, system call behavior, clock, random number, etc. I feel that ``all sources'' is perhaps an overclaim} \PDComment{We should clarify: as long as you model a source of nondet using \psharp, then you can capture it and control it (not automatically)}. We present two case studies of using \psharp to test production distributed systems inside Microsoft. Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time, as they did not have an effective way to reproduce it. \psharp uncovered the bug in a very small setting, which made it easy to examine traces, identify and fix the problem. \PDComment{we now have many bugs in Matt's system too.}

@@

\section{Testing Distributed Systems}
\label{sec:bg}

\input{background}
